\pkg{fitdistrplus}: An \proglang{R} Package for Distribution Fitting Methods

### 2013-03-28 JSS/fitdistrplus/paper2JSS.Rnw
last version before submit?
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/riskassessment@240 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Fitting distributions]{Fitting distributions to continuous non-censored data}\label{fitnoncenscont}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Choice of candidate distributions}
\label{Choice}
For illustrating the use of various functions of the \pkg{fitdistrplus} package 
with continuous non-censored data, we   first use
a data set named \code{groundbeef} which is included in our package.
This data set contains pointwise values of
serving sizes in grams, collected in a French survey, for ground beef patties consumed by 
children under 5 years old. It was used in a quantitative risk assessment 
published in the International Journal of Food Microbiology \cite{Delignette08}.
%%% R code set default options for all R schunks
<<par4vignette, echo=FALSE, results=hide>>=
options(digits=4, SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))) )
set.seed(1234)
@
%%% R code
<<datgroundbeef, echo=TRUE>>=
library(fitdistrplus)
data(groundbeef)
str(groundbeef)
@

@@

The data set named  \code{endosulfan} is now used to illustrate other features of 
the \pkg{fitdistrplus} package. This data set contains
acute toxicity values for the organochlorine pesticide endosulfan
(geometric mean of LC50 ou EC50 values in $\mu g.L^{-1}$),
tested on Australian and non-Australian laboratory-species 
(arthropods, fish or nonarthropod invertebrates, see \cite{Hose04}). 
<<datendosulfan, echo=TRUE>>=
data(endosulfan)
str(endosulfan)
@

@@

%%% R code
<<fitendo.echo, echo=TRUE, fig=FALSE>>=
ATV <-endosulfan$ATV
fendo.ln <- fitdist(ATV, "lnorm")
library(actuar)
fendo.ll <- fitdist(ATV, "llogis", start=list(shape=1, scale=500))
fendo.P <- fitdist(ATV, "pareto", start=list(shape=1, scale=500))
fendo.B <- fitdist(ATV, "burr", start=list(shape1=0.3, shape2=1, rate=1))
cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
        ylogscale = TRUE,legendtext = c("lognormal","loglogistic","Pareto","Burr"))
@
% qqcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale=TRUE, ylogscale=TRUE,
%       legendtext = c("lognormal","loglogistic","Pareto","Burr"))

@@

\begin{figure}[htb!]
  \centering
<<fitendo, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
        ylogscale = TRUE,legendtext = c("lognormal","loglogistic","Pareto","Burr"))
@
%qqcomp(list(fendo.ln,fendo.ll,fendo.P,fendo.B),xlogscale=TRUE,ylogscale=TRUE,
%       legendtext = c("lognormal","loglogistic","Pareto","Burr"))

@@

The computation of different goodness-of-fit statistics is proposed in the \pkg{fitdistrplus} package
in order to further compare fitted distributions.
The purpose of goodness-of-fit statistics aims to measure the distance between the
fitted parametric distribution and the empirical distribution: e.g. 
the distance between the fitted cumulative distribution function $F$ 
and the empirical distribution function $F_{n}$.
When fitting continuous distributions, three  goodness-of-fit statistics are classicaly considered: 
Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics \cite{Stephens86}.
Naming $x_{i}$ the $n$ observations of a continuous variable $X$ arranged in an ascending order,
Table \ref{tabKSCvMAD} gives the definition and the empirical estimate of the three considered 
goodness-of-fit statistics.

\begin{table}[htb!]
  \begin{center}
    \caption{Goodness-of-fit statistics as defined by Stephens \cite{Stephens86}.}
    \begin{tabular}{lll}
    \hline
      Statistic  & General formula & Computational formula\\
    \hline
      Kolmogorov-Smirnov   & $\sup|F_{n}(x) - F(x)|$ &
      $\max(D^{+},D^{-})$ with\\
      (KS)&& $D^{+}=\max\limits_{i=1,\dots,n}\left(\frac{i}{n} - F(x_{i})\right)
      ;D^{-}=\max\limits_{i=1,\dots,n}\left(F(x_{i})-\frac{i-1}{n}\right)$ \\
    \hline
      Cramer-von Mises  & n $\int_{-\infty}^{\infty}(F_{n}(x) - F(x))^2 dx$ &
        $\frac{1}{12n} + \sum\limits_{i=1}^n \left(F(x_{i})-\frac{2i-1}{2n} \right)^{2}$\\
      (CvM)&&\\
% cvm <- 1/(12*n) + sum( ( theop - (2 * seq(1:n) - 1)/(2 * n) )^2 )
    \hline
      Anderson-Darling  & n $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2}{F(x) (1 - F(x))} dx$
      & $-n -\frac{1}{n}\sum\limits_{i=1}^n \left((2i-1)(\log(F(x_{i}))+\log(1-F(x_{n+1-i})) )\right)$ \\
      (AD) &    & \\
% ad <- - n - mean( (2 * seq(1:n) - 1) * (log(theop) + log(1 - rev(theop))) )
    \hline
    \end{tabular}
    \label{tabKSCvMAD}
  \end{center}
\end{table}

They can be computed using the function \code{gofstat}
as defined by Stephens \cite{Stephens86}.

<<fendo.gof.print, echo=TRUE, fig=FALSE>>=
gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B))
@
As giving  more weight to distribution tails, 
Anderson-Darling statistics is of special interest when it matters to equally 
emphasize the tails as well as the main body of a distribution.
This is often the case in risk assessment, e.g. \cite{Cullen99,Vose10}. For this reason, 
this statistics is often used to select the best distribution among those fitted.
Nevertheless, this statistics should be used cautiously when comparing fits of various distributions. 
Keeping in mind that the weighting of each CDF quadratic difference 
depends on the parametric distribution in its definition (see Table \ref{tabKSCvMAD}), 
Anderson-Darling statistics computed for several distributions fitted on a same data set
are theoretically difficult to compare. Moreover, such a statistic, as Cramer-von Mises and
Kolmogorov-Smirnov ones, does not take into account the complexity of 
the model (i.e. parameter number). It is not
a problem when compared distributions are characterized by the same number of parameters, but
it could systematically promote the selection of the more complex distributions in the other case. 
Looking at classical penalized criteria based on the loglikehood (AIC, BIC) seems thus also interesting, especially
to discourage overfitting.

@@

\begin{table}[htb!]
  \begin{center}
    \caption{Modified Anderson-Darling statistics as defined by Luceno \cite{Luceno06}.}
    \begin{tabular}{lll}
    \hline
      Statistic  & General formula & Computational formula\\
    \hline
      Right-tail AD  & $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{1 - F(x)}  dx$
      & $\frac{n}{2} -2\sum_{i}F(x_{i}) -\frac{1}{n}\sum_{i}((2i-1)ln(1-F(x_{n+1-i})))$ \\
      (ADR) &  &\\
    \hline
      Left-tail AD  & $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(F(x))}  dx$
      & $-\frac{3n}{2} +2\sum_{i}F(x_{i}) -\frac{1}{n}\sum_{i}((2i-1)ln(F(x_{i})))$ \\
      (ADL) &  &\\
    \hline
      Right-tail AD  & $ad2r=\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(1 - F(x))^{2}}  dx$
      & $ad2r=2\sum_{i}ln(1-F(x_{i})) +\frac{1}{n}\sum_{i} \frac{2i-1}{1-F(x_{n+1-i})}$ \\
      2nd order (AD2R) &  &\\
    \hline
      Left-tail AD  & $ad2l=\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(F(x))^{2}}  dx$
      & $ad2l=2\sum_{i}ln(F(x_{i})) +\frac{1}{n}\sum_{i}\frac{2i-1}{F(x_{i})}$ \\
      2nd order (AD2L) &  &\\
    \hline
    AD 2nd order & $ad2r+ad2l$
      & $ad2r+ad2l$ \\
    (AD2) &  &\\
    \hline
    \end{tabular}
    \label{modifiedAD}
  \end{center}
\end{table}

@@

\begin{figure}[htb!]
  \centering
  %%% R code
<<mgegofcompplot, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L),
xlogscale = TRUE, ylogscale = TRUE,
main = "Fitting a lognormal distribution",
legendtext = c("MLE","Left-tail AD", "Left-tail AD 2nd order"),
xlegend = "bottomright")
@
  \caption{Comparison of a lognormal distribution fitted by MLE
  and by MGE using  two different goodness-of-fit distances : left-tail Anderson-Darling and left-tail Anderson Darling of seond order (example on the \code{endosulfan} data set)
  as provided by the \code{cdfcomp} function, with CDF values
  in a logscale to emphasize discrepancies on the left tail}
  \label{plotfitMGE}
\end{figure}

Comparing the $5\%$ percentiles (HC5) calculated using these three fits to the one calculated
from the MLE fit of the Burr distribution, we can observe, on this example,
that fitting the lognormal distribution by maximizing left tail 
Anderson-Darling distances of first or second order
enables to approach the value obtained by fitting the Burr distribution by MLE. 
%%% R code
<<quantilefitdist, echo=TRUE, fig=FALSE>>=
( HC5.estimates <- cbind(empirical = as.numeric(quantile(ATV, probs = 0.05)),
      Burr = as.numeric(quantile(fendo.B,probs = 0.05)$quantiles),
      lognormal_MLE = as.numeric(quantile(fendo.ln,probs = 0.05)$quantiles),
      lognormal_AD2 = as.numeric(quantile(fendo.ln.ADL,probs = 0.05)$quantiles),
      lognormal_AD2L = as.numeric(quantile(fendo.ln.AD2L,probs = 0.05)$quantiles)) )
@

@@

<<danish.mme.pareto, echo=TRUE, fig=FALSE>>=
library(actuar)
fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto",start=c(shape=10, scale=10), 
      lower=2+1e-6, upper=Inf)
memp <- function(x, order) ifelse(order == 1, mean(x), sum(x^order)/length(x))
fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", order=1:2, 
      memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), upper=c(Inf,Inf))
cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
        legend=c("Pareto MLE", "Pareto MME"), main="Fitting a Pareto distribution",
        xlogscale=TRUE, datapch=".")
gofstat(list(fdanish.ln.MLE, fdanish.P.MLE, fdanish.ln.MME, fdanish.P.MME))
@

\setkeys{Gin}{width=0.8\textwidth} %default

@@

\begin{figure}[htb!]
  \centering
  %%% R code
<<plotsalinity, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE, result=hide>>=
plotdistcens(salinity,Turnbull = FALSE)
@
  \caption{Simple plot of censored raw data (72-hour acute salinity tolerance of riverine macro-invertebrates from 
   the \code{salinity} data set) as ordered points 
   and intervals}
  \label{plotdistcens}
\end{figure}

@@

\setkeys{Gin}{width=0.5\textwidth} %default
\begin{figure}[htb!]
  \centering
  %%% R code
<<fitsalinitycdfcompplot, echo=FALSE, fig=TRUE, width=6, height=4, eps=FALSE>>=
cdfcompcens(list(fsal.ln, fsal.ll),
    legendtext=c("lognormal", "loglogistic "))
@
  \caption{Goodness-of-fit CDF plot for fits of a lognormal and a loglogistic distribution
  to censored data: LC50 values from 
   the \code{salinity} data set}
  \label{cdfcompcens}
\end{figure}

@@

\setkeys{Gin}{width=0.8\textwidth} %default

@@

\begin{figure}[htb]
  \centering
<<fittoxocarapoisnbinomplot, echo=FALSE, results=hide, fig=TRUE, width=6, height=4, eps=FALSE>>=
cdfcomp(list(ftoxo.P,ftoxo.nb),legendtext=c("Poisson", "negative binomial"))
@
  \caption{Comparison of the fits of a negative binomial and a Poisson distribution
   to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set)}
  \label{plotdiscfit}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Fitting distributions]{Fitting distributions to continuous non-censored data}\label{fitnoncenscont}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Choice of candidate distributions}
\label{Choice}
For illustrating the use of various functions of the \pkg{fitdistrplus} package 
with continuous non-censored data, we   first use
a data set named \code{groundbeef} which is included in our package.
This data set contains pointwise values of
serving sizes in grams, collected in a French survey, for ground beef patties consumed by 
children under 5 years old. It was used in a quantitative risk assessment 
published in the International Journal of Food Microbiology \cite{Delignette08}.
%%% R code set default options for all R schunks
<<par4vignette, echo=FALSE, results=hide>>=
options(digits=4, SweaveHooks=list(fig=function() par(mar=c(5.1, 4.1, 1.1, 2.1))))
set.seed(1234)
@
%%% R code
<<datgroundbeef, echo=TRUE>>=
library(fitdistrplus)
data(groundbeef)
str(groundbeef)
@

@@

The data set named  \code{endosulfan} is now used to illustrate other features of 
the \pkg{fitdistrplus} package. This data set contains
acute toxicity values for the organochlorine pesticide endosulfan
(geometric mean of LC50 ou EC50 values in $\mu g.L^{-1}$),
tested on Australian and non-Australian laboratory-species 
(arthropods, fish or nonarthropod invertebrates, see \cite{Hose04}). 
<<datendosulfan, echo=TRUE, eval=FALSE>>=
data(endosulfan)
str(endosulfan)
@
<<datendosulfan2, echo=FALSE, eval=TRUE>>=
data(endosulfan)
str(endosulfan, width = 70, strict.width = "cut")
@

@@

%%% R code
<<fitendo.echo, echo=TRUE, fig=FALSE>>=
ATV <-endosulfan$ATV
fendo.ln <- fitdist(ATV, "lnorm")
library(actuar)
fendo.ll <- fitdist(ATV, "llogis", start=list(shape=1, scale=500))
fendo.P <- fitdist(ATV, "pareto", start=list(shape=1, scale=500))
fendo.B <- fitdist(ATV, "burr", start=list(shape1=0.3, shape2=1, rate=1))
cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
      ylogscale = TRUE, legendtext = 
          c("lognormal", "loglogistic", "Pareto", "Burr"))
@
% qqcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale=TRUE, ylogscale=TRUE,
%       legendtext = c("lognormal","loglogistic","Pareto","Burr"))

@@

\begin{figure}[htb!]
  \centering
<<fitendo, echo=FALSE, fig=TRUE, width=4, height=4, eps=FALSE>>=
cdfcomp(list(fendo.ln, fendo.ll, fendo.P, fendo.B), xlogscale = TRUE,
        ylogscale = TRUE,legendtext = c("lognormal","loglogistic","Pareto","Burr"))
@
%qqcomp(list(fendo.ln,fendo.ll,fendo.P,fendo.B),xlogscale=TRUE,ylogscale=TRUE,
%       legendtext = c("lognormal","loglogistic","Pareto","Burr"))

@@

The computation of different goodness-of-fit statistics is proposed in the \pkg{fitdistrplus} package
in order to further compare fitted distributions.
The purpose of goodness-of-fit statistics aims to measure the distance between the
fitted parametric distribution and the empirical distribution: e.g. 
the distance between the fitted cumulative distribution function $F$ 
and the empirical distribution function $F_{n}$.
When fitting continuous distributions, three  goodness-of-fit statistics are classicaly considered: 
Cramer-von Mises, Kolmogorov-Smirnov and Anderson-Darling statistics \cite{Stephens86}.
Naming $x_{i}$ the $n$ observations of a continuous variable $X$ arranged in an ascending order,
Table \ref{tabKSCvMAD} gives the definition and the empirical estimate of the three considered 
goodness-of-fit statistics.
They can be computed using the function \code{gofstat}
as defined by Stephens \cite{Stephens86}.

\begin{table}[htb!]
  \begin{center}
    \caption{Goodness-of-fit statistics as defined by Stephens \cite{Stephens86}.}
    \begin{tabular}{lll}
    \hline
      Statistic  & General formula & Computational formula\\
    \hline
      Kolmogorov-Smirnov   & $\sup|F_{n}(x) - F(x)|$ &
      $\max(D^{+},D^{-})$ with\\
      (KS) & &  $D^{+}=\max\limits_{i=1,\dots,n}\left(\frac{i}{n} - F_i\right)$ \\
           & &  $D^{-}=\max\limits_{i=1,\dots,n}\left(F_{i}-\frac{i-1}{n}\right)$ \\
    \hline
      Cramer-von Mises  & n $\int_{-\infty}^{\infty}(F_{n}(x) - F(x))^2 dx$ &
        $\frac{1}{12n} + \sum\limits_{i=1}^n \left(F_i-\frac{2i-1}{2n} \right)^{2}$\\
      (CvM)&&\\
% cvm <- 1/(12*n) + sum( ( theop - (2 * seq(1:n) - 1)/(2 * n) )^2 )
    \hline
      Anderson-Darling  & n $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2}{F(x) (1 - F(x))} dx$
      & $-n -\frac{1}{n}\sum\limits_{i=1}^n (2i-1)\log(F_i(1-F_{n+1-i}))$ \\
      (AD) &    & \\
% ad <- - n - mean( (2 * seq(1:n) - 1) * (log(theop) + log(1 - rev(theop))) )
    \hline
     where $F_i\stackrel{\triangle}{=} F(x_i)$
    \end{tabular}
    \label{tabKSCvMAD}
  \end{center}
\end{table}

<<fendo.gof.print, echo=TRUE, fig=FALSE>>=
gofstat(list(fendo.ln, fendo.ll, fendo.P, fendo.B))
@

\newpage
As giving  more weight to distribution tails, 
Anderson-Darling statistics is of special interest when it matters to equally 
emphasize the tails as well as the main body of a distribution.
This is often the case in risk assessment, e.g. \cite{Cullen99,Vose10}. For this reason, 
this statistics is often used to select the best distribution among those fitted.
Nevertheless, this statistics should be used cautiously when comparing fits of various distributions. 
Keeping in mind that the weighting of each CDF quadratic difference 
depends on the parametric distribution in its definition (see Table \ref{tabKSCvMAD}), 
Anderson-Darling statistics computed for several distributions fitted on a same data set
are theoretically difficult to compare. Moreover, such a statistic, as Cramer-von Mises and
Kolmogorov-Smirnov ones, does not take into account the complexity of 
the model (i.e. parameter number). It is not
a problem when compared distributions are characterized by the same number of parameters, but
it could systematically promote the selection of the more complex distributions in the other case. 
Looking at classical penalized criteria based on the loglikehood (AIC, BIC) seems thus also interesting, especially
to discourage overfitting.

@@

\begin{table}[htb!]
  \begin{center}
    \caption{Modified Anderson-Darling statistics as defined by Luceno \cite{Luceno06}.}
    \begin{tabular}{lll}
    \hline
      Statistic  & General formula & Computational formula\\
    \hline
      Right-tail AD  & $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{1 - F(x)}  dx$
      & $\frac{n}{2} -2\sum\limits_{i=1}^nF_i -\frac{1}{n}\sum\limits_{i=1}^n(2i-1)ln(\overline F_{n+1-i})$ \\
      (ADR) &  &\\
    \hline
      Left-tail AD  & $\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(F(x))}  dx$
      & $-\frac{3n}{2} +2\sum\limits_{i=1}^nF_i -\frac{1}{n}\sum\limits_{i=1}^n(2i-1)ln(F_i)$ \\
      (ADL) &  &\\
    \hline
      Right-tail AD  & $ad2r=\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(1 - F(x))^{2}}  dx$
      & $ad2r=2\sum\limits_{i=1}^nln(\overline F_i) +\frac{1}{n}\sum\limits_{i=1}^n \frac{2i-1}{\overline F_{n+1-i}}$ \\
      2nd order (AD2R) &  &\\
    \hline
      Left-tail AD  & $ad2l=\int_{-\infty}^{\infty}\frac{(F_{n}(x) - F(x))^2 }{(F(x))^{2}}  dx$
      & $ad2l=2\sum\limits_{i=1}^nln(F_i) +\frac{1}{n}\sum\limits_{i=1}^n\frac{2i-1}{F_i}$ \\
      2nd order (AD2L) &  &\\
    \hline
    AD 2nd order & $ad2r+ad2l$
      & $ad2r+ad2l$ \\
    (AD2) &  &\\
    \hline
    where $F_i\stackrel{\triangle}{=} F(x_{i})$; & $\overline F_i\stackrel{\triangle}{=}1-F(x_{i})$
    \end{tabular}
    \label{modifiedAD}
  \end{center}
\end{table}

@@

\begin{figure}[htb!]
  \centering
  %%% R code
<<mgegofcompplot, echo=FALSE, fig=TRUE, width=4, height=4, eps=FALSE>>=
cdfcomp(list(fendo.ln, fendo.ln.ADL, fendo.ln.AD2L),
xlogscale = TRUE, ylogscale = TRUE,
main = "Fitting a lognormal distribution",
legendtext = c("MLE","Left-tail AD", "Left-tail AD 2nd order"),
xlegend = "bottomright")
@
  \caption{Comparison of a lognormal distribution fitted by MLE
  and by MGE using  two different goodness-of-fit distances : left-tail Anderson-Darling and left-tail Anderson Darling of seond order (example on the \code{endosulfan} data set)
  as provided by the \code{cdfcomp} function, with CDF values
  in a logscale to emphasize discrepancies on the left tail}
  \label{plotfitMGE}
\end{figure}

Comparing the $5\%$ percentiles (HC5) calculated using these three fits to the one calculated
from the MLE fit of the Burr distribution, we can observe, on this example,
that fitting the lognormal distribution by maximizing left tail 
Anderson-Darling distances of first or second order
enables to approach the value obtained by fitting the Burr distribution by MLE. 
%%% R code
<<quantilefitdist, echo=TRUE, fig=FALSE>>=
(HC5.estimates <- c(
empirical = as.numeric(quantile(ATV, probs=0.05)),
Burr = as.numeric(quantile(fendo.B, probs=0.05)$quantiles),
lognormal_MLE = as.numeric(quantile(fendo.ln, probs=0.05)$quantiles),
lognormal_AD2 = as.numeric(quantile(fendo.ln.ADL, probs=0.05)$quantiles),
lognormal_AD2L = as.numeric(quantile(fendo.ln.AD2L,probs=0.05)$quantiles)))
@

@@

<<danish.mme.pareto, echo=TRUE, fig=FALSE>>=
library(actuar)
fdanish.P.MLE <- fitdist(danishuni$Loss, "pareto", 
start=c(shape=10, scale=10), lower=2+1e-6, upper=Inf)
memp <- function(x, order) sum(x^order)/length(x)
fdanish.P.MME <- fitdist(danishuni$Loss, "pareto", method="mme", order=1:2, 
memp="memp", start=c(shape=10, scale=10), lower=c(2+1e-6,2+1e-6), 
upper=c(Inf,Inf))
cdfcomp(list(fdanish.P.MLE, fdanish.P.MME), 
legend=c("Pareto MLE", "Pareto MME"), main="Fitting a Pareto distribution",
xlogscale=TRUE, datapch=".")
gofstat(list(fdanish.ln.MLE, fdanish.P.MLE, fdanish.ln.MME, fdanish.P.MME))
@

\setkeys{Gin}{width=0.9\textwidth} %default

@@

\begin{figure}[htb!]
  \centering
  %%% R code
<<plotsalinity, echo=FALSE, fig=TRUE, width=6, height=6, eps=FALSE, result=hide>>=
plotdistcens(salinity,Turnbull = FALSE)
@
  \caption{Simple plot of censored raw data (72-hour acute salinity tolerance of riverine macro-invertebrates from 
   the \code{salinity} data set) as ordered points 
   and intervals}
  \label{plotdistcens}
\end{figure}

@@

\setkeys{Gin}{width=0.5\textwidth} %default
\begin{figure}[htb!]
  \centering
  %%% R code
<<fitsalinitycdfcompplot, echo=FALSE, fig=TRUE, width=6, height=6, eps=FALSE>>=
cdfcompcens(list(fsal.ln, fsal.ll),
    legendtext=c("lognormal", "loglogistic "))
@
  \caption{Goodness-of-fit CDF plot for fits of a lognormal and a loglogistic distribution
  to censored data: LC50 values from 
   the \code{salinity} data set}
  \label{cdfcompcens}
\end{figure}

@@

\setkeys{Gin}{width=0.9\textwidth} %default

@@

\begin{figure}[htb]
  \centering
<<fittoxocarapoisnbinomplot, echo=FALSE, results=hide, fig=TRUE, width=4, height=4, eps=FALSE>>=
cdfcomp(list(ftoxo.P,ftoxo.nb),legendtext=c("Poisson", "negative binomial"))
@
  \caption{Comparison of the fits of a negative binomial and a Poisson distribution
   to numbers of \emph{Toxocara cati} parasites from the \code{toxocara} data set)}
  \label{plotdiscfit}
\end{figure}
